---
title: 'ChemmineR: introductory session'
author: "Mik Black"
date: "25 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This material covers a VERY basic introduction to some of the functionality provided by the `ChemmineR` package for R.

It borrows heavily from: 

 - http://www.bioconductor.org/packages/devel/bioc/vignettes/ChemmineR/inst/doc/ChemmineR.html

### R libraries

The main R package used during this session is `ChemmineR`.  You can install this via:

```{r eval=FALSE}
## Source the biocLite.R installation script
source("http://bioconductor.org/biocLite.R") 

## Use biocLite to install the package
biocLite("ChemmineR") 
```

### Additional complication

Additional functionality can be added to the `ChemmineR` package through the OpenBabel C++ library, accessed via 
the `ChemmineOB` R package.  This provides extract format conversion and local peptide structure viewing capabilities.
As described in the manual, `ChemmineR` will automatically detect whether `ChemmineOB` is installed.  To get 
OpenBabel installed on my (Mac OS) laptop, I used the v2.3.1 zip file from:

http://openbabel.org/wiki/Category:Installation

and then installed the `ChemmineOB` package via:

```{r eval=FALSE}
source("http://bioconductor.org/biocLite.R")
biocLite("ChemmineOB")
```

### References

Papers relating to this package and the assocaited "Chemmine" web service:

 - Cao, Y., Charisi, A., Cheng, L.-C., Jiang, T., & Girke, T. (2008). ChemmineR: a compound mining framework for R. Bioinformatics (Oxford, England), 24(15), 1733–1734. http://doi.org/10.1093/bioinformatics/btn307

 - Backman, T. W. H., Cao, Y., & Girke, T. (2011). ChemMine tools: an online service for analyzing and clustering small molecules. Nucleic Acids Research, 39(suppl), W486–W491. http://doi.org/10.1093/nar/gkr320

Bioconductor page:

 - https://www.bioconductor.org/packages/release/bioc/html/ChemmineR.html

### Getting started

Load the ChemmineR library:

```{r}
library(ChemmineR)
```


